
The board of directors is:
\begin{itemize}
\item Aaron Krowne (PlanetMath Founder and President)
\item Joseph Corneli
\item Roger Lipsett
\item Raymond Puzio
\item Chi Woo
\end{itemize}

Board members are elected by voting members of PlanetMath, and serve for one
year. The board is responsible for deciding how PlanetMath's budget is spent,
including for hiring and overseeing staff or contractors; and for deciding and
maintaining site policies, including the appointment of site administrators.
The board is not compensated for this work, although board members can and do
at times serve as independent contractors with the organization. Our quarterly
board meetings that are open to the public; the dates of these meetings are
announced as news items on the website and on our mailing lists.

The mathematical content you find on the website is developed and maintained by
volunteers in our user community. Site users as a whole are responsible for
producing the content found on PlanetMath, and for reviewing and critiquing the
content other users have posted. A handful of users have special administrative
privileges. Site administrators can delete highly objectionable content and
suspend or remove a given user's posting privileges. Site administrators are
not compensated for their work.

Board members, administrators, and other active community members can be
reached through our mailing lists or the PlanetMath forums. The specific
policies of PlanetMath are published at:
\url{http://planetx.cc.vt.edu/AsteroidMeta/PlanetMath}.
Additional informal policy items of interest to site users are
available at: \url{http://planetmath.org/docs/}.
