\subsection*{Synopsis}

This brief document summarizes rules/regulations, conventions, and tips for using the PlanetMath forums.

\subsection*{Rules}

The rules for the PlanetMath forums are:

\begin{enumerate}
\item The forums are, for the most part, to be ``G-rated.'' So, no profanity, please. The only possible exception to the G-rated rule is the ``Humor'' forum, but please keep it to ``PG-13'' level here.
\item No flaming! How do you know if you are flaming? Basically, if you are arguing, not debating. Don't get emotional in the forums!
\item No spamming/commercial advertising! If you want to advertise on this site, buy Google ads. We keep the commercial and noncommercial stuff cleanly separated here.
\end{enumerate}

As you can see, the rules are few, and are pretty intuitive. However it is important to understand what makes these ``rules'' and not ``suggestions'': \textbf{if you break them, you are subject to enforcement} on this site. That means possibly getting locked out, having your account deleted and/or banned, and possibly losing control of any objects you own on the system. Enforcement is at the discretion of the administration!

\subsection*{Guidelines}

This section is for ``soft'' forum policy: guidelines, tips, and conventions. It is highly recommended that you follow them, but there is no administrative enforcement.

\begin{itemize}
\item When asking for help on mathematics, \emph{do not ask for the answer}. Nobody wants to do your work for you! Instead, state the problem and what you are confused about, and someone will almost certainly help you through it.
\item No ``shilling.'' Put another way, shamelessly self-(or otherwise) promoting advertising is frowned upon. However, it is ok to forward the bulletins of noncommercial math societies and the like.
\item Be polite, even when other people aren't! Be the more mature person! Messages that don't violate the rules will never go away; don't say anything you'll look back on and be ashamed of (of course, its ok and natural to be \emph{wrong}, but don't be \emph{mean}).
\end{itemize}

\subsection*{Updates}

Your help in developing this policy is welcome. Of course the administration has the last word on the rules, but feel free to add to the guidelines. Discussion of proposed changes to this document in the accompanying forum is recommended.
